Shapes analysis for time series

### 2024-10-25 NEURIPS24/appendix.tex
add last experiments

\section{Appendix for the experiment: Noise sensitivity}
This experiment evaluates the influence of noise on the learning of the reference sequence for TS-LDDMM and SRVF Kacher-mean, a subroutine of Shape-FPCA~\cite{wu2024shape}.

>>>>

\section{Appendix for the experiment: Noise sensitivity}
\label{appendix: noise sensitivity}
This experiment evaluates the influence of noise on the learning of the reference sequence for TS-LDDMM and SRVF Kacher-mean, a subroutine of Shape-FPCA~\cite{wu2024shape}.

### 2024-10-25 NEURIPS24/experiment.tex
add last experiments

    \item \textbf{Classification benchmark on regularly sampled datasets, \Cref{appendix: shape_classification}:} We compare performances of a kernel support vector machine (SVC) algorithm based on TS-LDDMM representation with 3 state-of-the-art classification methods from shape analysis on 15 shape-based datasets (7 univariates \& 8 multivariates). \textbf{Results} show that the TS-LDDMM-based method outperforms other methods (best performances over 13 datasets), making TS-LDDMM representation relevant for time series shape analysis.
\end{enumerate}

>>>>

    \item \textbf{Classification benchmark on regularly sampled datasets, \Cref{appendix: shape_classification}:} We compare performances of a kernel support vector machine (SVC) algorithm based on TS-LDDMM representation with 3 state-of-the-art classification methods from shape analysis on 15 shape-based datasets (7 univariates \& 8 multivariates). \textbf{Results} show that the TS-LDDMM-based method outperforms other methods (best performances over 13 datasets), making TS-LDDMM representation relevant for time series shape analysis.

    \item \textbf{Noise sensitivity for learning the reference graph, \Cref{appendix: noise sensitivity}:} We evaluate the noise sensitivity of TS-LDDMM and Shape-FPCA \cite{wu2024shape} for learning the reference graph on a synthetic dataset and for several levels of additive Gaussian noise. \textbf{Results} show that both methods are sensitive to noise. However, TS-LDDMM preserves the overall shape while shape-FPCA alters the shape depending on the noise level. 
\end{enumerate}

### 2024-10-25 NEURIPS24/neurips_2024_to_submit.tex
update
add last experiments

\title{Shapes analysis for time series.}

@@

\author{%
  Thibaut Germain \thanks{Corresponding author. Contact at \texttt{thibaut.germain@ens-paris-saclay.fr}} \\
  Centre Borelli, ENS Paris-Saclay\\
  4 av. des sciences, 91190 \\
  %\texttt{thibaut.germain@ens-paris-saclay.fr} \\
  % examples of more authors
  \And
  Samuel Gruffaz \\
  Centre Borelli, ENS Paris-Saclay\\
  4 av. des sciences, 91190 \\
  % Address \\
  %\texttt{samuel.gruffaz@ens-paris-saclay.fr} \\
  \AND
  Charles Truong \\
  Centre Borelli, ENS Paris-Saclay\\
  4 av. des sciences, 91190 \\
  %\texttt{charles.truong@ens-paris-saclay.fr} \\
  \And
  Laurent Oudre \\
  Centre Borelli, ENS Paris-Saclay\\
  4 av. des sciences, 91190 \\
  \And
  Alain Durmus \\
  CMAP, CNRS, Ecole polytechnique\\
  Institut Polytechnique de Paris\\
  91120 Palaiseau, France \\
}
\input{header.tex}
\input{def.tex}
\newcommand\sam[1]{\textcolor{blue}{#1}}
\newcommand\thi[1]{\textcolor{orange}{#1}}
\begin{document}

\maketitle
%In this paper we propose an unsupervised representation method (TS-LDDMM) to analyze inter-individual variability of time series dataset.
    %Inspired by the shape analysis literature, we extend Large Deformation Diffeomorphic Metric Mapping (LDDMM) to time series, considering the temporal evolution as their distinctive feature.  
    % Given a dataset of $N$ individual's time series $(s^j)_{j\in[N]}$,
    %  we learn a time series of reference $\mathbf{s}_0$ encoding the common part of the individuals as well as diffeomorphisms $(\phi^j)_{j\in[N]}$ encoding the specificities of each individual:
    %  $s^j$ is seen as the transformation of $\mathbf{s}_0$ by $\phi_j$.
    %  Then, the parameters $(\alpha_j)_{j\in[N]}$ encoding the diffeomorphisms $(\phi^j)_{j\in[N]}$ are used as the features representation related to each time series $s^j$.
    %At the crossroads between URL for time series and shape analysis. %the algorithm extends the Large Deformation Diffeomorphic Metric Mapping (LDDMM) to the case of time series. This extension is a nontrivial application of the LDDMM framework since it involves applying deformations to the graph of time series, which have more structure than curves due to the inclusion of the temporal dimension.

@@

\input{paper.tex}
\begin{ack}
  Use unnumbered first level headings for the acknowledgments. All acknowledgments
  go at the end of the paper before the list of references. Moreover, you are required to declare
  funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
  More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2024/PaperInformation/FundingDisclosure}.

  Do {\bf not} include this section in the anonymized submission, only in the final paper. You can use the \texttt{ack} environment provided in the style file to automatically hide this section in the anonymized submission.
  \end{ack}

>>>>

\title{Shapes analysis for time series}

@@

\author{%
  Thibaut Germain$^1$\thanks{Corresponding author. Contact at \texttt{thibaut.germain@ens-paris-saclay.fr}}\\
  Centre Borelli, ENS Paris-Saclay\\
  4 av. des sciences, 91190 \\
  %\texttt{thibaut.germain@ens-paris-saclay.fr} \\
  % examples of more authors
  \And
  Samuel Gruffaz$^1$\\
  Centre Borelli, ENS Paris-Saclay\\
  4 av. des sciences, 91190 \\
  % Address \\
  %\texttt{samuel.gruffaz@ens-paris-saclay.fr} \\
  \AND
  Charles Truong$^1$\\
  Centre Borelli, ENS Paris-Saclay\\
  4 av. des sciences, 91190 \\
  %\texttt{charles.truong@ens-paris-saclay.fr} \\
  \And
  Laurent Oudre$^1$\\
  Centre Borelli, ENS Paris-Saclay\\
  4 av. des sciences, 91190 \\
  \And
  Alain Durmus \\
  CMAP, CNRS, Ecole polytechnique\\
  Institut Polytechnique de Paris\\
  91120 Palaiseau, France \\
}

\input{header.tex}
\input{def.tex}

\newcommand\sam[1]{\textcolor{blue}{#1}}
\newcommand\thi[1]{\textcolor{orange}{#1}}

\begin{document}
\footnotetext[1]{Complete affiliation: Université Paris Saclay, Université Paris Cité, ENS Paris Saclay, CNRS, SSA, INSERM, Centre
Borelli, F-91190, Gif-sur-Yvette, France.}
\maketitle
%In this paper we propose an unsupervised representation method (TS-LDDMM) to analyze inter-individual variability of time series dataset.
    %Inspired by the shape analysis literature, we extend Large Deformation Diffeomorphic Metric Mapping (LDDMM) to time series, considering the temporal evolution as their distinctive feature.  
    % Given a dataset of $N$ individual's time series $(s^j)_{j\in[N]}$,
    %  we learn a time series of reference $\mathbf{s}_0$ encoding the common part of the individuals as well as diffeomorphisms $(\phi^j)_{j\in[N]}$ encoding the specificities of each individual:
    %  $s^j$ is seen as the transformation of $\mathbf{s}_0$ by $\phi_j$.
    %  Then, the parameters $(\alpha_j)_{j\in[N]}$ encoding the diffeomorphisms $(\phi^j)_{j\in[N]}$ are used as the features representation related to each time series $s^j$.
    %At the crossroads between URL for time series and shape analysis. %the algorithm extends the Large Deformation Diffeomorphic Metric Mapping (LDDMM) to the case of time series. This extension is a nontrivial application of the LDDMM framework since it involves applying deformations to the graph of time series, which have more structure than curves due to the inclusion of the temporal dimension.

@@

\input{paper.tex}
\begin{ack}
This work was supported by grants from Région Ile-de-France (DIM MathInnov).
Charles Truong is funded by the PhLAMES chair of ENS Paris-Saclay.
\end{ack}

### 2024-10-25 NEURIPS24/paper.tex
improve exp
add appendix
add last experiments

  \vspace{-1ex}
\section{Notations}
We denote by integer ranges by $[k:l]=\{k,\ldots,l\}\subset \mathcal{P}(\Zset)$ and $ [l]=[1:l]$ with $k,l\in \Nset$,
by $\rmC^m(\msi,\mse)$ the set of $m$-times continously differentiable function defined on an open set $\msu$ to a normed vector space $\mse$,
 by $||u||_\infty=\inf_{x\in \msu} |u(x)| $ for any bounded function $u:\msu \to \mse$,
and by $\Nset_{>0}$ is the set of positive integers. 
%We are in a sub class of curve quotiented by temporal reparametrisation since a time series' graph can be understood as the equivalent class of t ->(t,s(t))

@@

%\sam{améliorer transition}
\vspace{-1ex}
\paragraph{Loss}
\sam{Simplifier cette partie ?}
This section specifies the distance function $\scrl$ introduced in the loss function defined in \eqref{eq:relaxation}.

>>>>

  \vspace{-1ex}
\section{Notations}
We denote by integer ranges by $[k:l]=\{k,\ldots,l\}\subset \mathcal{P}(\Zset)$ and $ [l]=[1:l]$ with $k,l\in \Nset$,
by $\rmC^m(\msi,\mse)$ the set of $m$-times continously differentiable function defined on an open set $\msu$ to a normed vector space $\mse$,
 by $||u||_\infty=\sup_{x\in \msu} |u(x)| $ for any bounded function $u:\msu \to \mse$,
and by $\Nset_{>0}$ is the set of positive integers. 
%We are in a sub class of curve quotiented by temporal reparametrisation since a time series' graph can be understood as the equivalent class of t ->(t,s(t))

@@

%\sam{améliorer transition}
\vspace{-1ex}
\paragraph{Loss}
This section specifies the distance function $\scrl$ introduced in the loss function defined in \eqref{eq:relaxation}.
